The application layer is responsible for firing the application. In short, the job
of this layer is to parse the command line arguments and run the virtual machine.

This layer exists so that the main file is not populated and the fact that this layer could evolve by providing different ways to run
the program.

\begin{listing}[H]
\begin{minted}{cpp}
typedef enum : uint8_t {
    APPLICATION_FILE
} ApplicationType;
\end{minted}
\caption{Application types}
\label{ls:application_types}
\end{listing}

\autoref{ls:application_types} shows the supported application types. Additional types might be included in here, such as a prompt type or a string type.
At the moment, the supported application is a file application, meaning the program must come from a file path.
The application layer decides based on the command line arguments, the type of the application and performs the necessary actions accordingly.

This layer is also responsible for creating the \texttt{argv} parameter of the virtual machine. This parameter is a list of strings, so the C++ main arguments
\texttt{argc} and \texttt{argv} are transformed into a vector of strings and further passed to the virtual machine so it can inject it into the
main function call.

\section{Application class}

The application class is the implementation of the application layer, the implementation is very simple and can be seen in \autoref{ls:application_class}

\begin{listing}[H]
\begin{minted}{cpp}
class Application
{
    // Defines the application type described above.
    ApplicationType application_type;
    // Stores the vritual machine used by the application.
    VirtualMachine virtual_machine;
    // Stores the file name if the application type requires it.
    std::string file_name;
    // Stores the command line arguments.
    std::vector<std::string> argv;
    // Run the application based on an input string.
    void string(const std::string &string);
    public:
        // The constructor determines the type of the application
        // based on the command line arguments.
        Application(int argc, char *argv[]);
        // Starts (runs) the application.
        int start();
};
\end{minted}
\caption{Application class}
\label{ls:application_class}
\end{listing}
