All the Nuua virtual machine opcodes can be found in this list with a small comment on the right-hand side with the required operands of each
opcode. If the opcode has a result or destination, this is always the \emph{first} operand found in the opcode. For example the opcode
\texttt{OP\_MOVE RX RY} moves the value of \texttt{RY} to \texttt{RX}. As another example, the opcode \texttt{OP\_ADD\_INT RX RY RZ} performs
the addition of values \texttt{RY + RZ} and stores the result at \texttt{RX}. The following table explains the different operand types found.

\begin{itemize}
    \item \emph{RX, RY, RX, R1, R2, R3}: Represents the value of a register (they are found in the top frame).
    \item \emph{PX}: Represent the value of a register (they are found in the object)
    \item \emph{A}: Represents a literal value (normaly used for jumping offsets)
    \item \emph{G1}: Represents a global value (they are found in the global registers)
    \item \emph{C1}: Represents a constant value (they are found in the constant pool)\\
\end{itemize}

\begin{code}
\begin{minted}{cpp}
typedef enum : uint8_t {
    // Others
    OP_EXIT, // EXIT - - -

    /* Register manipulation */
    // Moves (copy) the value from RY to RX.
    OP_MOVE, // MOVE RX RY
    // Load a constant to the register.
    OP_LOAD_C, // OP_LOAD_C RX C1
    // Load a global register to a local register.
    OP_LOAD_G, // OP_LOAD_G RX G1
    // Set a global value.
    OP_SET_G, // OP_SET_G G1 RX

    /* Stack manipulation */
    // Push a value to the shared stack.
    OP_PUSH, // PUSH RX
    // Push a constant to the shared stack.
    OP_PUSH_C, // PUSH C1
    // Pops a value from the shared stack.
    OP_POP, // POP RX

    /* String releated */
    // Get the char from a string given its index.
    OP_SGET, // SGET RX RY RZ
    // Set a char of the string given its index.
    OP_SSET, // SSET RX RY RZ
    // Delete a given char in a string given its index.
    OP_SDELETE, // SDELETE RX RY

    /* List releated */
    // Push a value to a list.
    OP_LPUSH, // LPUSH RX RY
    // Push a constant to a list.
    OP_LPUSH_C, // LPUSH RX C1
    // Pop a value from a list.
    OP_LPOP, // LPOP RX
    // Get the value of a list given it's index.
    OP_LGET, // LGET RX RY RZ
    // Set the value of the list in the given index.
    OP_LSET, // LSET RX RY RZ
    // Delete a value of the list given the index.
    OP_LDELETE, // LDELETE RX RY

    /* Dictionary releated */
    // Get dictionary key given the integer index and the key order.
    OP_DKEY, // DKEY RX RY RZ
    // Get a value from the dictionary given the key.
    OP_DGET, // DGET RX RY RZ
    // Set the value of the given index key.
    OP_DSET, // DSET RX RY RZ
    // Delete a given key-value pair.
    OP_DDELETE, // DDELETE RX RY

    /* Function releated */
    // Call a function.
    OP_CALL, // CALL RX
    // Return to previous frame.
    OP_RETURN, // RETURN

    /* Object releated */
    // Get object property.
    OP_LPROP, // PROP RX PX
    // Set object property.
    OP_SPROP, // PROP PX RX

    /* Casting Operations */

    // Value casting
    OP_CAST_INT_FLOAT, // CAST_INT_FLOAT RX RY
    OP_CAST_INT_BOOL, // CAST_INT_BOOL RX RY
    OP_CAST_INT_STRING, // CAST_INT_STRING RX RY
    OP_CAST_FLOAT_INT, // CAST_FLOAT_INT RX RY
    OP_CAST_FLOAT_BOOL, // CAST_FLOAT_BOOL RX RY
    OP_CAST_FLOAT_STRING, // CAST_FLOAT_STRING RX RY
    OP_CAST_BOOL_INT, // CAST_BOOL_INT RX RY
    OP_CAST_BOOL_FLOAT, // CAST_BOOL_FLOAT RX RY
    OP_CAST_BOOL_STRING, // CAST_BOOL_STRING RX RY
    OP_CAST_LIST_STRING, // CAST_LIST_STRING RX RY
    OP_CAST_LIST_BOOL, // CAST_LIST_BOOL RX RY
    OP_CAST_LIST_INT, // CAST_LIST_INT RX RY
    OP_CAST_DICT_STRING, // CAST_DICT_STRING RX RY
    OP_CAST_DICT_BOOL, // CAST_DICT_BOOL RX RY
    OP_CAST_DICT_INT, // CAST_DICT_INT RX RY
    OP_CAST_STRING_BOOL, // CAST_STRING_BOOL RX RY
    OP_CAST_STRING_INT, // CAST_STRING_INT RX RY

    /* Unary Operations */

    // Negation
    OP_NEG_BOOL, // NEG_BOOL RX RY
    // Minus operations
    OP_MINUS_INT, // MINUS_INT RX RY
    OP_MINUS_FLOAT, // MINUS_FLOAT RX RY
    OP_MINUS_BOOL, // MINUS_BOOL RX RY
    // Plus operations
    OP_PLUS_INT, // PLUS_INT RX RY
    OP_PLUS_FLOAT, // PLUS_FLOAT RX RY
    OP_PLUS_BOOL, // PLUS_BOOL RX RY

    /* Extra binary but unary */
    // Increment register integer
    OP_IINC, // INC RX
    // Decrement register integer
    OP_IDEC, // DEC RX

    /* Binary Operations */

    // Addition
    OP_ADD_INT, // ADD_INT RX RY RZ
    OP_ADD_FLOAT, // ADD_FLOAT RX RY RZ
    OP_ADD_STRING, // ADD_STRING RX RY RZ
    OP_ADD_BOOL, // ADD_BOOL RX RY RZ
    OP_ADD_LIST, // ADD_LIST RX RY RZ
    OP_ADD_DICT, // ADD_DICT RX RY RZ
    // Substraction
    OP_SUB_INT, // SUB_INT RX RY RZ
    OP_SUB_FLOAT, // SUB_FLOAT RX RY RZ
    OP_SUB_BOOL, // SUB_BOOL RX RY RZ
    // Multiplication
    OP_MUL_INT, // MUL_INT RX RY RZ
    OP_MUL_FLOAT, // MUL_FLOAT RX RY RZ
    OP_MUL_BOOL, // MUL_BOOL RX RY RZ
    OP_MUL_INT_STRING, // MUL_INT_STRING RX RY RZ
    OP_MUL_STRING_INT, // MUL_STRING_INT RX RY RZ
    OP_MUL_INT_LIST, // MUL_INT_LIST RX RY RZ
    OP_MUL_LIST_INT, // MUL_LIST_INT RX RY RZ
    // Division
    OP_DIV_INT, // DIV_INT RX RY RZ
    OP_DIV_FLOAT, // DIV_FLOAT RX RY RZ
    OP_DIV_STRING_INT, // DIV_STRING_INT RX RY RZ
    OP_DIV_LIST_INT, // DIV_LIST_INT RX RY RZ
    // Equality
    OP_EQ_INT, // EQ_INT RX RY RZ
    OP_EQ_FLOAT, // EQ_FLOAT RX RY RZ
    OP_EQ_STRING, // EQ_STRING RX RY RZ
    OP_EQ_BOOL, // EQ_BOOL RX RY RZ
    OP_EQ_LIST, // EQ_LIST RX RY RZ
    OP_EQ_DICT, // EQ_DICT RX RY RZ
    // Not Equality
    OP_NEQ_INT, // NEQ_INT RX RY RZ
    OP_NEQ_FLOAT, // NEQ_FLOAT RX RY RZ
    OP_NEQ_STRING, // NEQ_STRING RX RY RZ
    OP_NEQ_BOOL, // NEQ_BOOL RX RY RZ
    OP_NEQ_LIST, // NEQ_LIST RX RY RZ
    OP_NEQ_DICT, // NEQ_DICT RX RY RZ
    // Higher than
    OP_HT_INT, // HT_INT RX RY RZ
    OP_HT_FLOAT, // HT_FLOAT RX RY RZ
    OP_HT_STRING, // HT_STRING RX RY RZ
    OP_HT_BOOL, // HT_BOOL RX RY RZ
    // Higher than or equal to
    OP_HTE_INT, // HTE_INT RX RY RZ
    OP_HTE_FLOAT, // HTE_FLOAT RX RY RZ
    OP_HTE_STRING, // HTE_STRING RX RY RZ
    OP_HTE_BOOL, // HTE_BOOL RX RY RZ
    // Lower than
    OP_LT_INT, // LT_INT RX RY RZ
    OP_LT_FLOAT, // LT_FLOAT RX RY RZ
    OP_LT_STRING, // LT_STRING RX RY RZ
    OP_LT_BOOL, // LT_BOOL RX RY RZ
    // Lower than or equal to
    OP_LTE_INT, // LTE_INT RX RY RZ
    OP_LTE_FLOAT, // LTE_FLOAT RX RY RZ
    OP_LTE_STRING, // LTE_STRING RX RY RZ
    OP_LTE_BOOL, // LTE_BOOL RX RY RZ

    /* Logical operations */
    // Logical or
    OP_OR, // OR RX RY RZ
    // Logical and
    OP_AND, // AND RX RY RZ

    /* Control flow (All relative jumps) */
    // Forward jump
    OP_FJUMP, // FJUMP A
    // Backward jump
    OP_BJUMP, // BJUMP A
    // Conditional forward jump
    OP_CFJUMP, // FJUMP A RX
    // Conditional backward jump
    OP_CBJUMP, // BJUMP A RX
    // Conditional forward negative jump
    OP_CFNJUMP, // FNJUMP A RX
    // Conditional backward negative jump
    OP_CBNJUMP, // BNJUMP A RX

    /* Slice and range */
    // String slice.
    OP_SSLICE, // SSLICE RX RY R1 R2 R3 (dest, target, start, end, step)
    // String slice to the end.
    OP_SSLICEE, // SSLICEE RX RY R1 R2 (dest, target, start, step)
    // List slice.
    OP_LSLICE, // LSLICE RX RY R1 R2 R3 (dest, target, start, end, step)
    // List slice to the end.
    OP_LSLICEE, // LSLICEE RX RY R1 R2 (dest, target, start, step)
    // Exclusive range.
    OP_RANGEE, // RANGEE RX R1 R2 (dest, start, end)
    // Inclusive range.
    OP_RANGEI, // RANGEI RX R1 R2 (dest, start, end)

    /* Utilities */
    // Print a register
    OP_PRINT, // PRINT RX
    // Print a constant
    OP_PRINT_C, // PRINT C1
} OpCode;
\end{minted}
\end{code}
